### 2013-01-06 documents/faqs_intro.Rnw
started report on prelim results and methods

\documentclass{article}
\usepackage{amsmath, amsthm, parskip, graphicx}

\frenchspacing

<<opts-set, message = FALSE, echo = FALSE>>=
require(knitr)
opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE,
               highlight = TRUE, background = 'white',
               fig.height = 3, fig.width = 4)
@

<<source-scripts>>=
source('../src/fish_plotting.r')
@

\title{Informal explination of study ``Ecotype detection in (geometric) morphometric data''}
\author{Peter D Smits \\
\and
Benjamin Frable}
\institute{Committee on Evolutionary Biology, University of Chicago \\
Department of Fisheries and Wildlife, Oregon State University}

\begin{document}
\maketitle

\section{Introduction}

\section{Materials and Methods}

\section{Results}

\section{Discussion}

\end{document}