### 2013-01-06 documents/faqs_intro.Rnw
more faq document.

\documentclass{article}
\usepackage{amsmath, amsthm, parskip, graphicx}

\frenchspacing

<<opts-set, message = FALSE, echo = FALSE>>=
require(knitr)
opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE,
               highlight = TRUE, background = 'white')
@

<<source-scripts, results = 'hide'>>=
source('../src/fish_plotting.r')
@

\title{Informal explination of study ``Ecotype detection in (geometric) morphometric data''}
\author{Peter D Smits \and Benjamin Frable}

\begin{document}
\maketitle

\section{Introduction}

\section{Materials and Methods}

\subsection{Fish}

\subsection{Morphometric analysis}

\subsection{Machine learning}

\subsubsection{Linear discriminate analysis}

\subsubsection{Multinomial logistic regression}

\section{Results}

\subsection{Morphometric analysis}

<<pls-fits, eval = FALSE>>=
gg.means
@
\begin{figure}[p]
  \centering
  \includegraphics{figure/pls-fits3}
  \caption{<+caption text+>}
  \label{fig:fish-fit}
\end{figure}

<<pca-plots, eval = FALSE>>=
gg.pca
@
\begin{figure}[p]
  \centering
  \includegraphics{figure/pca-plots3}
  \caption{<+caption text+>}
  \label{fig:fish-pca}
\end{figure}

\subsection{Machine learning}

\section{Discussion}

\end{document}