### 2013-03-29 documents/turtle_report.Rnw
turtle report filling out

\documentclass{article}
\usepackage{amsmath, amsthm}
\usepackage{graphicx, microtype, parskip}
\usepackage{rotating, longtable}
\usepackage[sort&compress]{natbib}

\frenchspacing

<<message = FALSE, echo = FALSE, warning = FALSE>>=
require(knitr)
require(xtable)
require(reshape2)
require(ggplot2)
require(GGally)

opts_chunk$set(warning = FALSE, echo = FALSE, message = FALSE,
               highlight = TRUE, background = 'white')
@

<<results = 'hide'>>=
source('../src/support_functions.r')
load('../src/ml_analysis.r')
@

\title{Progress report: \textit{Emys marmorata} sub-species identification and classification}

\begin{document}
\maketitle

\section{Methods}
No-free lunch theorem. Try lots of things because we don't understand everything.

\subsection{Unsupervised}
\subsubsection{Gap-based clustering}
Comparison of gap statistic results for partitioning around medoids (PAM) divisive clustering. Confidence intervals are determined via bootstrap. The higher the gap statistic, the better the clustering result.

\subsubsection{Evidence Accumulation Clustering}
Choosing an optimal number of partitions is hard, which is why gap-based cluster selection was used above. An alternative method is to look at the co-occurence frequency, that is how frequently any two samples occur in the same partition. Repeating this process over and over again creates the frequency, or ``vote'' for how the data set should be partitioned and  which specimens should be in the same cluster.

EAC was originally devised using \textit{k}-means clustering, but I've extended it to use PAM clustering instead. The hope is to determine underlying structure in the data given a wide enough partition range and a high enough number of iterations.

\subsection{Supervised}
\begin{itemize}
  \item multinomial logistic regression
  \item feed-forward neural networks
  \item random forests
\end{itemize}

\section{Preliminary results}

\section{Miscellaneous affairs}
\subsection{Evolution 2013}
Machine learning approaches for recognizing cyptic diversity: a case study using \texit{Emys marmorata}.

\subsection{Hinds grant}
Resubmit previous application with the results from this study in the prelim-results section?

Get network for one turtle? This should take long, just requires time.

\end{document}