### 2015-09-08 doc/second_tab.tex
added seven graph to document

% latex table generated in R 3.2.2 by xtable 1.7-4 package
% Tue Sep  8 13:30:59 2015
\begin{table}[ht]
\centering
\begin{tabular}{rrrr}
  \hline
 & npred & ii.auc & oo.auc \\ 
  \hline
RF &   11 & 0.999 & 1.000 \\ 
  LDA &   10 & 1.000 & 1.000 \\ 
  MnL &    4 & 1.000 & 0.999 \\ 
   \hline
\end{tabular}
\label{tab:second_res}
\end{table}