### 2015-09-10 doc/second_tab.tex

% latex table generated in R 3.2.2 by xtable 1.7-4 package
% Tue Sep  8 13:41:53 2015
\begin{table}[ht]
\centering
\begin{tabular}{rrrr}
  \hline
 & npred & ii.auc & oo.auc \\ 
  \hline
RF &   10 & 1.000 & 0.999 \\ 
  LDA &   10 & 1.000 & 1.000 \\ 
  MnL &    7 & 1.000 & 0.997 \\ 
   \hline
\end{tabular}
\label{tab:second_res}
\end{table}