### 2015-09-10 doc/tabs_2.tex

\documentclass{article}

\usepackage{amsmath, amsthm}
\usepackage{graphicx,hyperref}
\usepackage{microtype, parskip}
\usepackage{natbib}
\usepackage{lineno}
\usepackage[font=small]{caption}
\usepackage{subcaption, multirow, morefloats}
\usepackage{subcaption, wrapfig}
\usepackage{titlesec}
\usepackage[nottoc,numbib]{tocbibind}
\usepackage{authblk, attrib, fullpage}
\usepackage{lineno}

\frenchspacing

\setcounter{table}{1}

\begin{document}

\begin{table}
  \centering
  \caption{Results of bootstrap comparisons between the scheme with the highest mean AUC value and all other schemes, defined P(best - other \(>\) 0). An asterix indicates the best scheme. This was done for each of the three modeling techniques included in this study: random forest (RF), multinomial logistic regression (MLR), and linear discriminate analysis (LDA). Probabilities are the percent of comparisons that are greater than the observed difference in means.}
  \begin{tabular}{ l r r r }
    \hline
    Scheme & RF & MLR & LDA \\
    \hline
    \hline
    Morph 1 & \(\ast\) & \(\ast\) & 1 \\
    Morph 2 & 0.79 & 0.55 & 1 \\
    Mito 1 & 0.89 & 0.94 & 1 \\ 
    Mito 2 & 0.82 & 0.57 & \(\ast\) \\ 
    Mito 3 & 0.79 & 0.69 & 0.73 \\ 
    Nuclear & 0.79 & 0.96 & 0.96 \\ 
    \hline
  \end{tabular}
  \label{tab:gen_tests}
\end{table}

\end{document}