### 2015-09-10 doc/tabs_4.tex

\documentclass{article}

\usepackage{amsmath, amsthm}
\usepackage{graphicx,hyperref}
\usepackage{microtype, parskip}
\usepackage{natbib}
\usepackage{lineno}
\usepackage[font=small]{caption}
\usepackage{subcaption, multirow, morefloats}
\usepackage{subcaption, wrapfig}
\usepackage{titlesec}
\usepackage[nottoc,numbib]{tocbibind}
\usepackage{authblk, attrib, fullpage}
\usepackage{lineno}

\frenchspacing

\setcounter{table}{3}

\begin{document}

\begin{table}[ht]
  \centering
  \caption{Results from classification model estimates of the secondary, multi-species dataset. Models are random forest (RF), linear discriminate analysis (LDA), and multinomial logistic regression (MLR). Comparison of in-sample and out-of-sample AUC of the best performing model, along with the number of predictors. AUC values range between 0.5 (no better than random) and 1 (perfect classification).}
  \begin{tabular}{ l r r r }
    \hline
    & \# predictors & In-sample AUC & Out-of-sample AUC \\ 
    \hline
    \hline
    RF &   11 & 1.000 & 0.999 \\ 
    MLR &    6 & 1.000 & 0.998 \\ 
    LDA &   10 & 1.000 & 1.000 \\ 
    \hline
  \end{tabular}
  \label{tab:second_res}
\end{table}

\end{document}