### 2016-01-26 doc/second_tab.tex
changed variable number. had to updage labeller function because ggplot update

% latex table generated in R 3.2.3 by xtable 1.8-0 package
% Tue Jan 26 14:36:27 2016
\begin{table}[ht]
\centering
\begin{tabular}{rrrr}
  \hline
 & npred & ii.auc & oo.auc \\ 
  \hline
RF &   19 & 0.999 & 1.000 \\ 
  LDA &   11 & 1.000 & 1.000 \\ 
  MnL &    6 & 1.000 & 0.991 \\ 
   \hline
\end{tabular}
\label{tab:second_res}
\end{table}