### 2016-03-03 doc/second_tab.tex
updated figures

% latex table generated in R 3.2.3 by xtable 1.8-0 package
% Thu Mar  3 14:40:31 2016
\begin{table}[ht]
\centering
\begin{tabular}{rrrr}
  \hline
 & npred & ii.auc & oo.auc \\ 
  \hline
RF &   22 & 1.000 & 1.000 \\ 
  LDA &   10 & 1.000 & 1.000 \\ 
  MnL &    3 & 1.000 & 1.000 \\ 
   \hline
\end{tabular}
\label{tab:second_res}
\end{table}